### 2024-07-12 actuarial.qmd
20240712 - gambler's fallacy

$$
\scriptsize
\text{Violence risk} = \beta \cdot \text{conduct disorder} + \beta \cdot \text{substance use} + \beta \cdot \text{suspended from school} + \beta \cdot \text{childhood aggression} + ...
$$ {#eq-actuarial}

>>>>

$$
\scriptsize
\text{violence risk} = \beta \cdot \text{conduct disorder} + \beta \cdot \text{substance use} + \beta \cdot \text{suspended from school} + \beta \cdot \text{childhood aggression} + ...
$$ {#eq-actuarial}

### 2024-07-12 cognitive-bias.qmd
20240711 - copyediting
20240712 - gambler's fallacy

```{r}

```

@@

The gambler's fallacy occurs due to an erroneous belief in the law of small numbers.
The law of large numbers is a mathematical theorem that the average of a sufficiently large number of independent observations converges to the true value.
For instance, if you flip a fair coin 1 million times, it is likely to land heads-up ~50% of the time.
The law of *small* numbers (aka hasty generalization), by contrast, is an erroneous belief that small samples are representative of the populations from which they were drawn.
For instance, if you flip a coin 10 times, belief in the law of small numbers would lead one to believe that the coin will flip heads-up exactly 5 times out of 10.
However, in reality, the chance is less than 1 in 4 (24.6%) that exactly 5 of 10 coin flips turn up heads:

@@

@@

In conclusion, there are many [heuristics](#sec-heuristics), [cognitive bias](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
It is prudent to be aware of these common biases and to work to counteract them.
For instance, look for information that challenges or disconfirms your beliefs, and work to incorporate this information into your beliefs.
Do your best to pursue observation, data analysis, and data interpretation with an open mind.
You never know what important information you might discover if you go in with an open mind.
Pay attention to [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation) when considering whether a player's performance may [regress to the mean](#sec-fallaciesRegression).
If the strong performances appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
There is a lot of luck in fantasy football.
Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.

>>>>

```{r}
library("tidyverse")
```

@@

The gambler's fallacy occurs due to an erroneous belief in the law of small numbers.
The law of large numbers is a mathematical theorem that the average of a sufficiently large number of independent observations converges to the true value.
For instance, if you flip a fair coin 1 million times, it is likely to land heads-up ~50% of the time.
The law of *small* numbers (aka hasty generalization), by contrast, is an erroneous belief that small samples are representative of the populations from which they were drawn.
For instance, if you flip a coin 10 times, belief in the law of small numbers would lead one to believe that the coin will flip heads-up exactly 5 times out of 10.
However, in reality, the chance is less than 1 in 4 (24.6%) that exactly 5 of 10 coin flips turn up heads, as calculated below and as depicted in @fig-simulationOf10CoinFlips:

@@

The `dbinom()` function in R provides the density of a binomial distribution.
A binomial distribution is the probability of a particular number of successes (e.g., coins flipping heads-up) given a certain number of independent trials.

```{r}
#| fig-cap: "Histogram of number of coins that flip heads-up in a simulation of 10 coin flips (with 100,000 replications)."
#| label: fig-simulationOf10CoinFlips
#| code-fold: true

set.seed(52242)

numHeads <- rbinom(
  n = 100000,
  size = 10,
  prob = .5
)

simulationOfFlipping10Coins <- data.frame(
  numHeads = numHeads
)

simulationOfFlipping10Coins <- simulationOfFlipping10Coins %>% 
  mutate(
    highlight = ifelse(numHeads == 5, "yes", "no")
  )

ggplot(
  data = simulationOfFlipping10Coins,
  mapping = aes(
    x = numHeads,
    fill = highlight)
) +
  geom_histogram(
    color = "#000000",
    bins = 11
  ) +
  scale_x_continuous(
    breaks = 0:10
  ) +
  scale_fill_manual(
    values = c(
      "yes" = "tomato",
      "no" = "gray")
  ) +
  theme_classic() +
  theme(legend.position = "none")
```

Although 5 is the modal count of coins that flip heads-up out of 10 flips (i.e., it was more common than any other number), it is less common than the aggregate probability of flipping any number of heads besides 5.
The probability of getting any other number of coin flips turning up heads (other than 5) is:

```{r}
dbinom(
  x = c(0:4, 6:10),
  size = 10,
  prob = 0.5
) %>% sum()
```

@@

In conclusion, there are many [heuristics](#sec-heuristics), [cognitive bias](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
It is prudent to be aware of these common biases and to work to counteract them.
For instance, look for information that challenges or disconfirms your beliefs, and work to incorporate this information into your beliefs.
Do your best to pursue observation, data analysis, and data interpretation with an open mind.
You never know what important information you might discover if you go in with an open mind.
Pay attention to [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation) when considering whether a player's performance may [regress to the mean](#sec-fallaciesRegression).
If the strong performances appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
There is considerable luck in fantasy football.
Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.
